Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-11-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Fix subproof below Axiom 2 in page 18
Make minor improvements to proofs
Improve acknowledgements.

%implication for equivalence in essence: 
%\begin{small}
\begin{prooftree}
\AXC{Axiom \ref{A:A3}} \dashedLine
\UIC{$P(G)$}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\wedge_I$}
\BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\imp_E$}
\AXC{Axiom \ref{A:A2}} 
\BIC{$P(\varphi)$}
\AXC{$\subproof_{11}$} \dashedLine 
\UIC{$  G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$ \all \varphi. (P(\varphi)\biimp \nec \varphi(\gamma))$}  \RightLabel{$\all_E$}
\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
\UIC{$[\subproof_{14}] \quad \nec  \forall y. (G(y)\imp \varphi(y))\imp \nec \varphi(\gamma)$}
\end{prooftree}
%\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_{13}$ and $\subproof_{14}$} \dashedLine \RightLabel{$\wedge_I$}
\UIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
\UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y))) $}
\dottedLine \RightLabel{D\ref{A:D2}}
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine 
\AXC{$\subproof_{11}$} \dashedLine
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
\BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$}
   % \AXC{Axiom T} \dashedLine
    %\UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
\UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
\UIC{$ \all \varphi.(\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)) $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
%\end{tiny}
\end{proof}

@@

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\biimp_E$}
\AXC{Axiom \ref{A:A3}} \RightLabel{}
\UIC{$P(G)$} \RightLabel{$\biimp_E$}
\BIC{$\nec G(\gamma)$}
\AXC{$ $} \RightLabel{1}
\UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^1$}
\UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec (G(\gamma)\imp \ex x.G(x))$} \doubleLine  \RightLabel{(Axiom K)}
\UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^2$}
\UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
\end{prooftree}
%\end{scriptsize}

@@

\paragraph{Acknowledgements:} We thank Padre Edvaldo (from Piracicaba, Brazil) for assisting Bruno and his family in a moment of need. His help triggered the beginning of this work in 2013. We are grateful to Christoph Benzm\"uller for collaborations on several related works. This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).

>>>>

%implication for equivalence in essence: 
%\begin{small}
\begin{prooftree}
\AXC{Axiom \ref{A:A3}} \dashedLine
\UIC{$P(G)$}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\wedge_I$}
\BIC{$P(G)\wedge \nec  \forall y. (G(y)\imp \varphi(y))$}\RightLabel{$\imp_E$}
\AXC{Axiom \ref{A:A2}} \dashedLine
\UIC{\ldots\footnote{The formulas are ommitted here due to lack of space.}} \doubleLine \RightLabel{$\all_E$ x 2}
\UIC{\ldots}
\BIC{$P(\varphi)$}
\AXC{$\subproof_{11}$} \dashedLine 
\UIC{$  G(\gamma)$} \dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$ \all \varphi. (P(\varphi)\biimp \nec \varphi(\gamma))$}  \RightLabel{$\all_E$}
\UIC{$ P(\varphi)\biimp \nec \varphi(\gamma)$}  \RightLabel{$\biimp_E$}
\BIC{$ \nec \varphi(\gamma)$} \RightLabel{$\imp_I^5$}
\UIC{$[\subproof_{14}] \quad \nec  \forall y. (G(y)\imp \varphi(y))\imp \nec \varphi(\gamma)$}
\end{prooftree}
%\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_{13}$}
\AXC{$\subproof_{14}$} \RightLabel{$\wedge_I$}
\BIC{$\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y)) $} \RightLabel{$\forall_I$}
\UIC{$\forall \varphi. (\nec \varphi(\gamma)\biimp \nec  \forall y. (G(y)\imp \varphi(y))) $}
\dottedLine \RightLabel{D\ref{A:D2}}
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine 
\AXC{$\subproof_{11}$} \dashedLine
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{A:D1}}
\UIC{$\forall \varphi.(P(\varphi) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \biimp \nec E(\gamma)$}\RightLabel{$\biimp_E$}
\BIC{$\nec E(\gamma)$}\RightLabel{$\nec_E$}
   % \AXC{Axiom T} \dashedLine
    %\UIC{$\nec E(\gamma) \imp E(\gamma)$} \RightLabel{$\imp_E$}
\UIC{$ E(\gamma)$} \dottedLine \RightLabel{D\ref{A:D3}}
\UIC{$ \all \varphi.(\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)) $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
%\end{tiny}
\end{proof}

@@

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\biimp_E$}
\AXC{Axiom \ref{A:A3}} \dashedLine \RightLabel{}
\UIC{$P(G)$} \RightLabel{$\biimp_E$}
\BIC{$\nec G(\gamma)$}
\AXC{$ $} \RightLabel{1}
\UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^1$}
\UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec (G(\gamma)\imp \ex x.G(x))$} \doubleLine  \RightLabel{(Axiom K)}
\UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^2$}
\UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
\end{prooftree}
%\end{scriptsize}

@@

\paragraph{Acknowledgements:} We thank Padre Edvaldo (from Piracicaba, Brazil) for assisting Bruno and his family in a moment of need. Gratitude for his assistance is what triggered the production of an initial draft of this article in early 2013. We are thankful to Christoph Benzm\"uller for collaborations on several related articles. This work was supported by an Stipendium of the \"Osterreichische Akademie der Wissenschaften (APART).
